---
title: "Modeling increases in global wild fish catch 1950-2012"
description: |
  Here I estimate parameters with nonlinear least squares for the increase in global wild fish catch 1950-2012. 
author:
  - name: Peyton Moore
    url: {}
date: 03-14-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE, warning = FALSE, include = TRUE)

library(tidyverse)
library(janitor)
library(here)
library(boot)
library(gt)
library(patchwork)
library(broom)
library(nlstools)

```


## Read in data


Source: Global wild fish catch and aquaculture production, compiled by Earth Policy Institute with 1950-2010 from U.N. Food and Agriculture Organization (FAO), Global Capture Production and Global Aquaculture Production, electronic databases, at www.fao.org/fishery/topic/16140/en.

```{r}
fish_catch <- read_csv(here("data", "fish_catch.csv"), skip = 2) %>% 
  drop_na() %>% 
  clean_names() %>% 
  mutate(year = as.numeric(year), 
         wild_catch = as.numeric(wild_catch), 
         farmed_fish = as.numeric(farmed_fish), 
         total_fish_production = as.numeric(total_fish_production))

#exploratory graph

ggplot(data = fish_catch)+
  geom_point(aes(x = year, 
             y = wild_catch), 
             color = "blue")+
  geom_point(aes(x = year, 
             y = farmed_fish), 
             color = "red")+
  geom_point(aes(x = year, 
             y = total_fish_production), 
             color = "yellow")

ggplot(data = fish_catch)+
  geom_point(aes(x = year, 
             y = log(wild_catch)), 
             color = "blue")

```
Wild catch looks logistic 
Production of farmed fish resembles an exponential function
total fish production is linear-ish 

Logistic growth equation:

$P(t)=\frac{K}{1+Ae^{-kt}}$, where

- $K$ is the carrying capacity
- $A$ is $\frac{K-P_0}{P_0}$
- $k$ is the growth rate constant




```{r}

fish_years <- fish_catch %>% 
  mutate(model_yr = year-1950)


```

```{r}
df_nls <- nls(wild_catch ~ K/(1 + A*exp(-r*model_yr)),
              data = fish_years,
              start = list(K = 180, A = 17, r = 0.17),
              trace = TRUE
              )


summary(df_nls)

model_out <- broom::tidy(df_nls)

p_predict <- predict(df_nls)

df_complete <- data.frame(fish_years, p_predict)

```

Parameters:

K = 1.003e+02
A = 4.316e+00
r = 6.989e-02


```{r}

ggplot(data = df_complete, aes(x = year, y = wild_catch)) +
  geom_point(color = "darkorange",
             alpha = 0.6) +
  geom_line(aes(x = year, y = p_predict)) +
  theme_bw()+
  labs(x = "Year",
       y = "Wild Catch in Million Tons", 
       title = "Wild Catch from 1950-2012")
```


