---
title: "Cetacean abundance in California"
description: |
  Here is a course project where I showcase my skills in spatial analysis of 'raster()' data. 
author:
  - name: Peyton Moore
    url: {}
date: 03-14-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE, warning = FALSE, include = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(sf)
library(tmap)
library(gstat)
library(stars)
library(raster)
library(rnaturalearth)
```

### Data

Kaschner, K., Rius-Barile, J., Kesner-Reyes, K., Garilao, C., Kullander, S., Rees, T., & Froese, R. (2016). AquaMaps: Predicted range maps for aquatic species. www.aquamaps.org


### Reading in and exploring the Rasters
```{r}

# read in cetacean data 
my_files <- list.files(path = here("data", "ca_cetaceans"), full.names = TRUE)

#create a stack of all the different species 
cetacean_data <- raster::stack(my_files)

#plot it to see whats going on - ick
plot(cetacean_data)

#turn it into a data frame to see what is going on
cetacean_df <- rasterToPoints(cetacean_data) %>% 
  as.data.frame()

```

### Create a pressence vs. absence function 

```{r}
#create a function to determine if there is a probability of seeing each species in each cell, threshold of 0.6 

presence_of_cetacean <- function(x, thresh = 0.6){
  y <- ifelse(x >= thresh, 1, NA)
  return(y)
}

# apply that 
cet_rich <- calc(cetacean_data, fun = presence_of_cetacean)

plot(cet_rich, col = 'green4')


```

### Create data frame 

- Use `rasterToPoints()` 
- Create a species column with `pivot_longer()`
- Summarize to find the total number of species present in each cell

```{r}

richness_df <- rasterToPoints(cet_rich) %>% 
  as.data.frame() %>% 
  pivot_longer(cols = 3:37, names_to = "species", values_to = "presence") %>% 
  group_by(x, y) %>% 
  summarize(richness = sum(presence, na.rm = TRUE))
```

### Read in World Map From `rnatualearth`

```{r}
#world data 

nat_earth <- ne_countries(scale = "medium", returnclass = "sf")


st_crs(nat_earth)

nat_earth <- st_transform(nat_earth, st_crs(cetacean_data))
 
st_crs(nat_earth)

plot(nat_earth)


```


### Use `ggplot` with `geom_raster()` to create a finalized figure of cetacean richness along the California Coast



```{r, include=TRUE}
ggplot()+
  geom_raster(data = richness_df, aes(x = x, y = y, fill= richness))+
  geom_sf(data = nat_earth) +
  coord_sf(expand = 0, xlim = c(-125, -115), ylim =c(32,38))+
  scale_fill_gradient(low = 'darkslategray1', high = 'turquoise4') +
  theme_minimal()+
  labs(fill = "Cetacean Species Richness", 
       x = "Longitude", 
       y = "Latitude", 
       title = "Cetacean Species Richness Along the California Coast")

```
**Figure 1:** Species richness for 35 cetacean species is shown along the California Coast, with light blues indicating lower species richness, and darker blues indicating higher species richness. This value of richness is extrapolated from probability data, where a threshold of 0.6 was used. Species with a probability of presence in the area of greater than 0.6 were considered to be present, and species with a probability lower than 0.6 were considered to be absent. 
