---
title: "Oceanography parameters at LTER sites in Santa Barbara"
description: |
  Here I create a simple dendrogram using clustering analysis to represent different oceanographic parameters in Santa Barbara. 
author:
  - name: Peyton Moore
    url: {}
date: 03-14-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(here)
library(boot)
library(gt)
library(patchwork)
library(broom)
library(nlstools)

# Packages for cluster analysis:
library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
```


### Read in the data 

```{r}
sbc_lter <- read_csv(here("data", "sbc_lter_registered_stream_chemistry.csv"))%>% 
  mutate(tpc_uM = replace(tpc_uM, 
                        tpc_uM %in% c("-999.0", "-999", "NaN"),
                        NA)) %>% 
   mutate(tdp_uM = replace(tdp_uM, 
                        tdp_uM %in% c("-999.0", "-999", "NaN"),
                        NA)) %>% 
   mutate(nh4_uM = replace(nh4_uM, 
                        nh4_uM %in% c("-999.0", "-999", "NaN"),
                        NA)) %>% 
   mutate(no3_uM = replace(no3_uM, 
                        no3_uM %in% c("-999.0", "-999", "NaN"),
                        NA)) %>% 
   mutate(po4_uM = replace(po4_uM, 
                        po4_uM %in% c("-999.0", "-999", "NaN"),
                        NA)) %>% 
   mutate(tdn_uM = replace(tpc_uM, 
                        tdn_uM %in% c("-999.0", "-999", "NaN"),
                        NA)) %>% 
   mutate(tpn_uM = replace(tpn_uM, 
                        tpn_uM %in% c("-999.0", "-999", "NaN"),
                        NA)) %>% 
   mutate(tpp_uM = replace(tpc_uM, 
                        tpp_uM %in% c("-999.0", "-999", "NaN"),
                        NA)) %>% 
   mutate(tss_mgperLiter = replace(tss_mgperLiter, 
                        tss_mgperLiter %in% c("-999.0", "-999", "NaN"),
                        NA)) %>%
  mutate(spec_cond_uSpercm = replace(spec_cond_uSpercm, 
                        spec_cond_uSpercm %in% c("-999.0", "-999", "NaN"),
                        NA)) 
# create a summary by site 
lter_sum <- sbc_lter %>% 
  group_by(site_code) %>% 
  summarize(nh4_uM = mean(nh4_uM, na.rm = TRUE), 
            no3_uM = mean(no3_uM, na.rm = TRUE), 
            po4_uM = mean(po4_uM, na.rm = TRUE), 
            tdn_uM = mean(tdn_uM, na.rm = TRUE), 
            tdp_uM = mean(tdp_uM, na.rm = TRUE), 
            tpc_uM = mean(tpc_uM, na.rm = TRUE), 
            tpn_uM = mean(tpn_uM, na.rm = TRUE), 
            tpp_uM = mean(tpp_uM, na.rm = TRUE), 
            tss_mgperLiter = mean(tss_mgperLiter, na.rm = TRUE), 
            spec_cond_uSpercm = mean(spec_cond_uSpercm, na.rm = TRUE)) %>% 
  drop_na()
#scale the data 
lter_summary_scaled <- lter_sum %>% 
  select(2:11) %>% 
  scale()
#add row names back in 
rownames(lter_summary_scaled) <- lter_sum$site_code

# calc euclidean distance 

euc_distance <- dist(lter_summary_scaled, method = "euclidean")


view(euc_distance)
```

DATA & METADATA SOURCE: 
SBC LTER: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000
Creators: Santa Barbara Coastal LTER, & Melack, John M
Citation: Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative. https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174. 


### Do hierarchical clustering to create a dendrogram 

```{r}
cluster_lter <- hclust(euc_distance, method = "complete")



ggdendrogram(cluster_lter, 
             rotate = "TRUE")+
  theme_minimal()+
  coord_flip()+
  labs(x = "Site Code", 
       y = "", 
       title = "Dendrogram of LTER sites based on oceanography paramaters")
```




